Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2014-03-20 ms.tex
more on absolute zero point

\subsection{Zero-Point}
\label{sec:zp}
Although we wish to isolate and model thermal emission from Galactic dust, the
\PLANCK~maps contain additional components on large angular scales. At each 
frequency, there can exist an overall, constant offset that must be subtracted 
to set the zero level of Galactic dust. Additionally, faint residuals of the 
Solar dipole remain at low frequencies. At 353 GHz, 545 GHz, 857 GHz, we have 
adopted the zero-point offsets of \cite{planckdust} Table 1, which were derived
 by correlating \PLANCK~857 GHz versus H\,\textsc{i} at low column density, and
then by correlating the other \PLANCK~channels against the corrected 857 GHz 
map. We have not altered the \verb|i100| zero-point in any way. We have also 
removed the Solar dipole residual according to \cite{planckdust}, by 
subtracting a dipole pattern oriented toward 
$(l, b) = (263.99^{\circ}, 48.26^{\circ})$ and scaled according to their 
Table 1.

@@

\subsection{Finite Bandpass Correction}
The thermal dust emission model of $\S$\ref{sec:modeling} predicts the 
intensity $M_{\nu}$ in MJy/sr for any single frequency $\nu$. In practice, 
however, we wish to constrain our model using measurements in the broad 
\IRAS/\PLANCK~bandpasses, typically hundreds of GHz wide. We have therefore 
computed dimensionless bandpass correction factors $b_{\nu_i}$ such that 
$M_{\nu_i}b_{\nu_i}$ is the predicted observation for the \PLANCK~or \IRAS~band
 of interest, in MJy/sr. Here $\nu_i$ represents the $i^{th}$ of the five 
nominal band center frequencies for which we have broadband measurements, 
$\nu_i \in$ \{217, 353, 545, 857, 3000\} GHz. The shape of model spectrum 
$M_{\nu}$ is only dependent upon one parameter: the hot dust temperature $T_2$.
 We have used the \PLANCK~transmission curves provided in the file 
\verb|HFI_RIMO_R1.10.fits| \citep{planckresponse}. For 3000 GHz, we have used 
the \IRAS~bandpass, following the treatment of \cite{planckdust}. We note that 
it is not clear which bandpass should properly be ascribed to the \IRAS/DIRBE 
composite 100$\mu$m maps under consideration.

>>>>

\subsection{Zero-Point}
\label{sec:zp}

% for quantitative assessment of the zero-point difference due to
% flattening of 857 GHz vs. HI, see 
% ~/Desktop/dust/planck_hi/piecewise.pdf (on my Mac)

% -> at low HI, slope decreases by a factor of 1.88
%    the resulting difference in 857 GHz zero points is 0.373 MJy/sr

Although we wish to isolate and model thermal emission from Galactic dust, the
\PLANCK~maps contain additional components on large angular scales. At each 
frequency, there can exist an overall, constant offset that must be subtracted 
to set the zero level of Galactic dust. Additionally, faint residuals of the 
Solar dipole remain at low frequencies. We will address these issues by
separately solving two sub-problems: first, fixing the absolute zero level 
relative to external data, and second fitting per-band relative offsets 
and low order corrections by correlating the \PLANCK~bands against one another.
%this last sentence needs lots of work

\subsubsection{Absolute Zero Level}
In \cite{planckdust}, the absolute zero level of infrared thermal dust emission
was set by requiring that \PLANCK~infrared emission tends to zero when 
H\,\textsc{i} is zero, assuming a linear correlation between these two 
measurements at low column density. However, this approach is somewhat
unsatisfying in that there appear to be different slopes of \PLANCK~857 GHz 
versus H\,\textsc{i} for different ranges of H\,\textsc{i} intensity. In 
particular, \PLANCK~857 GHz appears to ``flatten out'' at very low 
H\,\textsc{i}. More quantitavely, we have found using the LAB H\,\textsc{i} 
data \citep{lab} for $-72$$<$$v_{LSR}$$<$$+25$ that the best-fit slope for 
H\,\textsc{i}$<$70 K km/s is a factor of $\sim$1.9 lower than the best fit 
slope for 110 K km/s $<$H\,\textsc{i}$<$200 K km/s, and as a result the implied
zero-point offsets for \PLANCK~857 GHz differ by $\sim$0.37 MJy/sr.

% CITE KALBERLA LAB PAPER !!!!!

Because of these ambiguities in the relationship between 857 GHz and 
H\,\textsc{i} emission, we decided to instead constrain the 857 GHz zero level
by comparison to the FDS99 857 GHz prediction. This renders the \PLANCK~857 GHz
absolute zero level tied largely to the reliable FIRAS zero point, and to some 
lesser extent to H\,\textsc{i} through the \verb|i100| zero point. 
Specifically, we fit a linear model to the FDS99 predicted 857 GHz values as a 
function of \PLANCK~857 GHz. We used a version of the \PLANCK~857 GHz map with 
zodiacal light and point sources removed and smoothed to 1$^{\circ}$ 
resolution, which we will refer to as $\mathcal{I}_{857}$. For this analysis, 
we considered $\mathcal{I}_{857}$ to be the independent variable, as it has 
much higher S/N than the FDS99 prediction, henceforward referred to as 
$\mathcal{F}_{857}$. Note that $\mathcal{F}_{857}$ is not a monochromatic 
prediction, as we have incorporated the finite bandpass correction factors 
described in $\S$\ref{sec:bpcorr}, using the FDS99 temperature map to determine
the dust spectrum shape. We then rebin to $N_{side}$=64 and restrict to pixels 
with $\mathcal{I}_{857}$$<$2.15 MJy/sr. Since \PLANCK~857 GHz smoothed to 
degree resolution has very high S/N, we can safely perform such a cut on 
$\mathcal{I}_{857}$. Figure XX shows a scatterplot of $\mathcal{I}_{857}$ 
versus $\mathcal{F}_{857}$, with a moving median and linear fit overplotted.
The linear fit was performed with uniform weights and iterative outlier
rejection. It is encouraging that the slope is quite close to unity, and 
the offset near zero.

The formal statistical errors on these parameters are small. Undoubtedly,
just as in the case of performing the regression versus  H\,\textsc{i}, 
systematic issues will dominate the uncertainty in the derived zero point
offset. Specifically, the likely culprits are imperfect zodiacal light 
corrections and the FDS99 temperature map. To quantify the level of systematic
error

% actually evaluate the statistical errors !!!!

% in previous sentence fill in specifics of slope and offset

% sentence saying slope is close to unity, and relationship indeed looks linear

% sentence or two explaining split into quadrants, differences in slope
% and offset probably due to zodiacal light and/or issues with FDS T2 map
% say how this quadrants exercise (via rms) allows us to define 857 GHz
% uncertainty

% what HI intensity does the 2.15 MJy/sr 857 GHz correspond to ?
%  ---> pretty high actually

% mention use of bandpass correction factors when doing Planck vs. prediction
% regressions

\subsubsection{Relative Zero Level}

At 353 GHz, 545 GHz, 857 GHz, we have 
adopted the zero-point offsets of \cite{planckdust} Table 1, which were derived
 by correlating \PLANCK~857 GHz versus H\,\textsc{i} at low column density, and
then by correlating the other \PLANCK~channels against the corrected 857 GHz 
map. We have not altered the \verb|i100| zero-point in any way. We have also 
removed the Solar dipole residual according to \cite{planckdust}, by 
subtracting a dipole pattern oriented toward 
$(l, b) = (263.99^{\circ}, 48.26^{\circ})$ and scaled according to their 
Table 1.

@@

\subsection{Finite Bandpass Correction}
\label{sec:bpcorr}
The thermal dust emission model of $\S$\ref{sec:modeling} predicts the 
intensity $M_{\nu}$ in MJy/sr for any single frequency $\nu$. In practice, 
however, we wish to constrain our model using measurements in the broad 
\IRAS/\PLANCK~bandpasses, typically hundreds of GHz wide. We have therefore 
computed dimensionless bandpass correction factors $b_{\nu_i}$ such that 
$M_{\nu_i}b_{\nu_i}$ is the predicted observation for the \PLANCK~or \IRAS~band
 of interest, in MJy/sr. Here $\nu_i$ represents the $i^{th}$ of the five 
nominal band center frequencies for which we have broadband measurements, 
$\nu_i \in$ \{217, 353, 545, 857, 3000\} GHz. The shape of model spectrum 
$M_{\nu}$ is only dependent upon one parameter: the hot dust temperature $T_2$.
 We have used the \PLANCK~transmission curves provided in the file 
\verb|HFI_RIMO_R1.10.fits| \citep{planckresponse}. For 3000 GHz, we have used 
the \IRAS~bandpass, following the treatment of \cite{planckdust}. We note that 
it is not clear which bandpass should properly be ascribed to the \IRAS/DIRBE 
composite 100$\mu$m maps under consideration.
